---
title: 'Practical Machine Learning Project : Prediction Assignment Writeup'
author: "Chien-Hua Wang"
date: "March 23, 2019"
output:
  word_document: default
  html_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```
#Practical machine Learning Project: Prediction Assignment Writeup

##Overview
###This document is the final project of the Peer Assessment project from Coursera’s course Practical Machine Learning, as part of the Specialization in Data Science. It was built up in RStudio, using its knitr functions, meant to be published in html format.
###This analysis meant to be the basis for the course quiz and a prediction assignment writeup. The main goal of the project is to predict the manner in which 6 participants performed some exercise as described below. This is the “class” variable in the training set. The machine learning algorithm described here is applied to the 20 test cases available in the test data and the predictions are submitted in appropriate format to the Course Project Prediction Quiz for automated grading.

##Background
###Using devices such as Jawbone Up, Nike FuelBand, and Fitbit it is now possible to collect a large amount of data about personal activity relatively inexpensively. These type of devices are part of the quantified self movement - a group of enthusiasts who take measurements about themselves regularly to improve their health, to find patterns in their behavior, or because they are tech geeks. One thing that people regularly do is quantify how much of a particular activity they do, but they rarely quantify how well they do it. In this project, your goal will be to use data from accelerometers on the belt, forearm, arm, and dumbell of 6 participants. They were asked to perform barbell lifts correctly and incorrectly in 5 different ways. More information is available from the website here: http://groupware.les.inf.puc-rio.br/har (see the section on the Weight Lifting Exercise Dataset).

##Library Import
```{r, echo=TRUE}
library(knitr)
library(lattice)
library(ggplot2)
library(caret)
library(rpart)
library(rpart.plot)
library(rattle)
library(randomForest)
library(corrplot)
```

##Data Loading and Exploratory Analysis
```{r, echo=TRUE}
# set the URL for the download
UrlTrain <- "http://d396qusza40orc.cloudfront.net/predmachlearn/pml-training.csv"
UrlTest  <- "http://d396qusza40orc.cloudfront.net/predmachlearn/pml-testing.csv"

# download the datasets
training <- read.csv(url(UrlTrain))
testing  <- read.csv(url(UrlTest))

# create a partition with the training dataset 
inTrain  <- createDataPartition(training$classe, p=0.7, list=FALSE)
TrainSet <- training[inTrain, ]
TestSet  <- training[-inTrain, ]
dim(TrainSet)

#Checking Missing Values
colnames(TrainSet)[sapply(TrainSet,function(x) sum(is.na(x))) > 0]
#Delete these columns which have extreme low variance
#Using Near Zero Variance (NZV)
NZV <- nearZeroVar(TrainSet)
TrainSet <- TrainSet[, -NZV]
TestSet  <- TestSet[, -NZV]
dim(TrainSet)
dim(TestSet)
```
```{r,echo=TRUE}
#remove variables that are mostly NA
AllNA    <- sapply(TrainSet, function(x) mean(is.na(x))) > 0.95
TrainSet <- TrainSet[, AllNA==FALSE]
TestSet  <- TestSet[, AllNA==FALSE]
dim(TrainSet)
dim(TestSet)

#remove identification between 1 and 5 columns
TrainSet <- TrainSet[, -(1:5)]
TestSet  <- TestSet[, -(1:5)]
dim(TrainSet)
dim(TestSet)
```
##Correlation Analysis
```{r, echo=TRUE}
corMatrix = cor(TrainSet[,-54])
corrplot(corMatrix, order="FPC", method="color", type="lower",tl.cex = 0.8, tl.col = rgb(0,0,0))
```
###The highly correlated variables are shown in dark colors in the graph above. To make an evem more compact analysis, a PCA (Principal Components Analysis) could be performed as pre-processing step to the datasets. Nevertheless, as the correlations are quite few, this step will not be applied for this assignment.

##Prediction Model Building
###Three methods will be applied to model the regressions (in the Train dataset) and the best one (with higher accuracy when applied to the Test dataset) will be used for the quiz predictions. The methods are: Random Forests, Decision Tree and Generalized Boosted Model, as described below.
###A Confusion Matrix is plotted at the end of each analysis to better visualize the accuracy of the models.

##Random Forest
```{r,echo=TRUE}
set.seed(12345)
controlRF = trainControl(method="cv", number=3, verboseIter = FALSE)
modFitRandomForest = train(classe~., data = TrainSet, method='rf',trControl=controlRF)
modFitRandomForest$finalModel
```

```{r,echo=TRUE}
# prediction on Test dataset
predictRandForest <- predict(modFitRandomForest, newdata=TestSet)
confMatRandForest <- confusionMatrix(predictRandForest, TestSet$classe)
confMatRandForest
```
```{r, echo=TRUE}
# plot matrix results
plot(confMatRandForest$table, col = confMatRandForest$byClass, 
     main = paste("Random Forest - Accuracy =",
      round(confMatRandForest$overall['Accuracy'], 3)))
```

##Decision Tree
```{r,echo=TRUE}
# model fit
set.seed(12345)
modFitDecTree <- rpart(classe ~ ., data=TrainSet, method="class")
prp(modFitDecTree)
```
```{r, echo=TRUE}
# prediction on Test dataset
predictDecTree <- predict(modFitDecTree, newdata=TestSet, type="class")
confMatDecTree <- confusionMatrix(predictDecTree, TestSet$classe)
confMatDecTree
```
```{r,echo=TRUE}
# plot matrix results
plot(confMatDecTree$table, col = confMatDecTree$byClass, 
     main = paste("Decision Tree - Accuracy =",
                  round(confMatDecTree$overall['Accuracy'], 3)))
```
##Generalized Boosted Model
```{r,echo=TRUE}
# model fit
set.seed(12345)
controlGBM <- trainControl(method = "repeatedcv", number = 5, repeats = 1)
modFitGBM  <- train(classe ~ ., data=TrainSet, method = "gbm",
                    trControl = controlGBM, verbose = FALSE)
modFitGBM$finalModel
```
```{r,echo=TRUE}
# prediction on Test dataset
predictGBM <- predict(modFitGBM, newdata=TestSet)
confMatGBM <- confusionMatrix(predictGBM, TestSet$classe)
confMatGBM
```
```{r,echo=TRUE}
# plot matrix results
plot(confMatGBM$table, col = confMatGBM$byClass, 
     main = paste("GBM - Accuracy =", round(confMatGBM$overall['Accuracy'], 3)))
```

#Test Dataset Accuracy comparison
```{r,echo=TRUE}
results = list(RF=confMatRandForest$overall[1], DT=confMatDecTree$overall[1], GBM=confMatGBM$overall[1])
results
```
###Above the table, the Random Forest model will be applied to predict the 20 quiz results (testing dataset) as shown below.
```{r, echo=TRUE}
predictTEST <- predict(modFitRandomForest, newdata=testing)
predictTEST
```



